---
title: "How to use"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(histoc)
```
 
The `histoc` assembles functions for histocompatibility testing in the context of kidney transplantation.   

Package main functions allows to simulate kidney allocation systems for the distribution of deceased donors' grafts.

Currently, it is possible to simulate allocation rules implemented in Portugal (PT model), in countries within EuroTransplant^[EuroTransplant. ETKAS and ESP; 2020] (ET model), in the United Kingdom^[NHSBT, KAP - Kidney Allocation Policy. Published 2016] (UK model), and a system previously suggested by Lima *et al.*^[Lima BA, Mendes M, Alves H. Kidney Transplant allocation in Portugal. Port J Nephrol Hypert. 2013;27(4):313-316] (Lima's model), using functions: `pts()`, `et()`, `uk()` and `lima()`, respectively.

Each one of these functions have as arguments a data frame with transplant candidates' clinical and demographic characteristics, a data frame with candidates' HLA antibodies and data for a potential donor.

For all the algorithms, a virtual crossmatch between the donor and transplant candidates is performed (`xmatch()`). And, only those candidates with a negative crossmatch and ABO compatible can opt to the donor's kidney.

Results are presented as `data.table` objects due to its high computation performance.

## PT model

```{r}
pts()
```

By default 2 candidates are selected for each donor, although we can define the number of candidates to be selected.

Portuguese rules on allocation of kidneys from deceased donor (PT model) are based on a scoring system
that takes in consideration: HLA mismatches between donor and transplant candidate, level of immunization of the candidate, time on dialysis, and age difference between donor and transplant candidate. Total scores for donor-recipient pairs are given by the column `ptsPT`. Also, hipersensitized candidates (`hiper()`) (cPRA >85%, `HI`) are prioritized and after that
all candidates are ordered by their corresponding score.

## ET model

```{r}
et()
```

A simplified version of EuroTransplant Kidney Allocation System (ETKAS)^[EuroTransplant. ETKAS and ESP; 2020] for first time kidney only candidates with more than 18 years old and not having donated one of their own kidneys can be simulated through `et()`.

In this simulation for each donor, transplant candidates are sorted as :

1. Senior Program (65+ years of old candidates when the donor has 65+ years) `SP`;
1. Acceptable Mismatch Program (candidates with a cPRA >85% and without HLA antibodies against HLA's donor) `AM`;
1. 000 HLA mismatches (candidates without HLA mismatches with the donor) `mmHLA`;
1. ETKAS points `pointsET`.

Final points for each eligible candidate are obtained from the sum of HLA points (`et_mmHLA()`), dialysis (`et_dialysis()`) points and MMP points (`et_mmp()`).


## UK model

```{r}
uk()
```


United Kingdom (UK) deceased donor kidney allocation for transplantation is firstly based on the definition of two ranked Tiers of eligible candidates for a donor (`Tier`):

1. Tier A – patients with match score = 10 or `cPRA` = 100% or time on `dialysis` >7 years;
1. Tier B – all other eligible patients.

Within Tier A, transplant candidates are ordered by matchability and time on dialysis. Transplant candidates within Tier B are prioritized according to a point-based system computed with 7 elements:

1. Matchability (`matchability`)
1. Time on dialysis (`dialysis`)
1. Donor-recipient risk index combinations (`ric()`)
1. HLA match and age combined (`age` and `mmHLA`)
1. Donor-recipient age difference (`age` and `donor_age`)
1. Total HLA mismatch (`mmHLa`)
1. Blood group match (`abo_uk()`)

This function simulates the allocation of kidneys to a candidates' waiting list for kidney-only transplants and do not take in consideration any geographical criteria.

## Lima's model

```{r}
lima()
```

Within this model, a color prioritization (`cp()`) of all waiting list transplant candidates is established.

Transplant candidates are classified according to their clinical urgency (red color), and regarding their time on `dialysis` and cPRA value `cPRA`. With an orange color are marked those patients with a `cPRA` >85% or with a time on `dialysis` higher than waiting time 3rd quartile. As yellow are classified the patients with a `cPRA` >50% or with a time on `dialysis` higher than waiting time median. And, as green are classified all the rest.

Within each color group candidates are ordered by `mmHLA` (ascendant) and time on `dialisys` (descendant).

Also, candidates are allocated to donors within the same age group (old to old program) (`SP`), mimicking EuroTransplant senior program (`sp()`).


