---
title: "Candidate's selection"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(histoc)
```

## Candidates’ selection for a pool of donors

We can also simulate the selection of wait list candidates for a pool of donors, according to a given model (or algorithm). 
The function `donor_recipient_pairs()` allow us to compute all possible donor-recipient pairs according to any of the [previously described](how_to.html) kidney allocation algorithms. 

Here an example where we use a pool of 70 donors ([donors](../reference/donors.html) data frame available in the package) to select from a wait list of 500 transplant [candidates](../reference/candidates.html) (data frame is also available in the package)


```{r}
# compute all possible donor-recipient pairs for each one of the donors according to ‘lima’ algorithm
all_pairs <- donor_recipient_pairs(df.donors = donors,
                      df.candidates = candidates,
                      df.abs = cabs,
                      algorithm = lima,
                      n = 0,
                      check.validity = FALSE)

# for loop to select 2 available candidates for a pool of donors
used.candidates <- NULL
result <- NULL
for(i in 1: length(all_pairs)){
  tmp <- all_pairs[[i]][!ID %in% used.candidates][1:2,]

  result <- data.table::rbindlist(list(result, tmp))
  used.candidates <- c(used.candidates, tmp$ID)
}

# as a result we have all the selected donor-recipient pairs for a given pool of donors
result
```

Now we can also compute a new column for the estimated 5-year event (mortality or graft failure) probability as described by Molnar *et al.*^[Predictive Score for Posttransplantation Outcomes. Molnar M, Nguyen D, Chen Y, et al. Transplantation. 2017 Jun;101(6):1353-1364] and available from the application [TxScore](https://balima.shinyapps.io/scoreTx/), with the function `txscore()`:

```{r}
result[!is.na(ID),][ ,
                     txs := txscore(recipient.age = age,
                                    recipient.dialysis = dialysis,
                                    donor.age = donor_age,
                                    mmHLA_A = mmA,
                                    mmHLA_B = mmB,
                                    mmHLA_DR = mmDR)$prob5y,
                     by = 'ID'][] 
```


## Input data

Input data used with this package's functions, regarding either candidates or donors information, when provided by the user must have the exact same format as the example data available.   Moreover, [{simK}](https://github.com/txopen/simK) package allows to generate synthetic data both for candidates and donors that can be used with {histoc}.  



